
\section{Introduction}

In the previous chapters, we have discussed visualisation and its role in bridging the gap between data and understanding. We have applied centrality metrics to a chemical network to tell us what species are of importance and experimented in getting machine learning models to learn the chemical structure of the species in a mechanism. This final research chapter provides a brief overview of current mechanism reduction techniques, whilst providing two novel alternatives to aid the process.

Science often deals with the problem of understanding complexity. This may be accomplished through organisation and partitioning, for example, the learning of a new skill through chunking, or the parallelisation of a large mathematical problem. In cases where such methods fail, we are forced to `disregard' complexity. It is common to approximate an atom as a sphere or the value $\pi$ as 3 with little consequence to the overall result of a calculation. The process of lumping has long been used to replace a complex, changing process (e.g Quantum Mechanics or Boundary Layer Fluid Dynamics) with a simpler constant process, \citep{approx}. In such cases, an approximate analysis may be far more useful than a lengthy exact solution, or none at all. 

Similar problems of complexity can also be seen within the chemistry of the atmosphere. An example is seen within the Master Chemical Mechanism\footnote{Version 3.3.1 .} (MCM), \citep{mcm}, this contains 1228 \ch{RO2} reactions. If written explicitly all \ce{RO2-RO2} interactions would result in a total of 1507984 reactions. Instead, the MCM overcomes this problem by creating a \ch{ro2} pool, with which all \ch{RO2} species react. This results in a mechanism which preserves the quality of science with only 0.000814 of the total possible \ch{ro2-ro2} reactions.

However even with such simplifications, atmospheric chemical mechanisms have been increasing in size over the last 10 years, \autoref{fig:mcmgoogleREF}REFC1fig. With the ability to automate their construction, mechanisms with species numbers of the millions become possible. Although the existence of more-explicit mechanisms may improve the quality of science produced, they can cause problems for efficient computation, diagnosis and analysis. This chapter shall look at two methods in which we may simplify a mechanism by grouping similar species together. These are through the use of species lifetime (\autoref{sec:lifetime}) and graph-based clustering (\autoref{sec:graphreduction}).


